---
title: "Practical Machine Learning Course Project"
author: "Vijay Kumar"
date: "March 10, 2016"
output: html_document
---


Packages Required
```{r}
library(dplyr)
library(randomForest)
library(caret)

```

Loading the training and test data
```{r}
setwd("C:/Users/vkumar16/Desktop/R Data_Assg/Project_ML/practicalmachinelearning")
training<- read.csv("pml-training.csv", header = TRUE)
testing<- read.csv("pml-testing.csv", header = TRUE)
```

Summarizing the data
```{r}
summary(training)
```

Identifying & Converting training and test data into “NA”, “” and “#DIV/0!” as NA strings
```{r}
training<- read.csv("pml-training.csv", header = TRUE, na.strings= c("NA", "", "#DIV/0!"))
testing<- read.csv("pml-testing.csv", header = TRUE, na.strings= c("NA", "", "#DIV/0!"))

```

Splitting into training data into training and test
```{r}
intrain<- createDataPartition(y= training$classe, p=0.7, list= FALSE)
traindata <- training[intrain,]
testdata <- training[-intrain,]
```

Remove X, username and cvdt_timestamp as predictors from the training subset
```{r}
traindata<- traindata[,-grep("X|user_name|cvtd_timestamp", names(traindata))]
```

Remove Near Zero predictors
```{r}
traindata<- traindata[, -nearZeroVar(traindata)]
```

NA Treatment
```{r}
NAPercent<- colSums(is.na(traindata))/nrow(traindata)
NAMajority <- names(NAPercent[NAPercent > 0.70])
NAtoRemove<- which(NAPercent > 0.70)
traindata <- traindata[,-NAtoRemove]
```

List of predictor variables
```{r}
Var<- names(traindata)
ModelVar<- Var[-grep("classe", Var)]
ModelVar
```

Random Forest Model based on ModelVar predicting classe
```{r}
set.seed(83)
datatomodel<- traindata[, Var]
modelfit <- randomForest(classe ~., data=datatomodel, type="class")

```

Insample Error & Confusion Matrix
```{r}
predtrain <- predict(modelfit,newdata=traindata)
confusionMatrix(predtrain,traindata$classe)$table
```

Outofsample Error & Confusion Matrix
```{r}
predtest <- predict(modelfit,newdata=testdata)
confusionMatrix(predtest,testdata$classe)
```